---
title: "Project Modeling"
author: "STOR 320.002 Group 2"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(rvest)
library(noncensus)
library(readr)
library(knitr)

#Import modeling
library(modelr)
library(broom)

# Import Data Below
EFI = read_csv("Clean_EFI.csv") 
EFI2 = EFI
write.csv(EFI2, "Clean_EFI.csv", row.names = F)
```

```{r}
MAE.func <- function(x) {
  mean(abs(x))
}

RMSE.func <- function(x) {
  sqrt(mean(x^2))
}
```

```{r}
# Clean EFI Data, Remove NA's, and Find Correlations with EFS
EFI_DATA <- rename(EFI, EFS = `2019_Score`) %>%
  select(Country, EFS, Tariff_Rate:Public_Debt_percent_of_GDP) %>%
  filter(Country != "Venezuela") %>%  # Exclude Venezuela
  filter(!is.na(EFS),
         !is.na(GDP_per_Capita_PPP),
         !is.na(Tariff_Rate),
         !is.na(Inflation_rate),
         !is.na(Tax_Burden_percent_of_GDP),
         !is.na(FDI_Inflow_Millions),
         !is.na(Corporate_Tax_Rate),
         !is.na(Unemployment_rate))

#Finding the variables with the higheset correlation with EFS(2019_score)
correlations = EFI_DATA %>%
  select(EFS, Tariff_Rate:Public_Debt_percent_of_GDP) %>%
  cor(use="complete.obs") %>%
  .[1,]

# Correlations with EFS
correlations
```

```{r}
#SPLIT EFI INTO TEST AND TRAIN DATASETS
set.seed(123)

EFI_DATA = EFI_DATA %>%
  mutate(SET = sample(x=c("TRAIN", "TEST"), 
         size=n(),
         replace=T, 
         prob=c(0.8, 0.2)))

TRAIN = EFI_DATA %>% filter(SET=="TRAIN")
TEST = EFI_DATA %>% filter(SET=="TEST")
```


```{r}
# Model 0: All 14 variables
lm_full <- lm(EFS ~ Govt_Expenditure_percent_of_GDP + Tax_Burden_percent_of_GDP + 
              GDP_Growth_Rate + FDI_Inflow_Millions + GDP_Billions_PPP +
              Population_Millions + Corporate_Tax_Rate + Unemployment_rate +
              Tax_Burden_percent_of_GDP + Govt_Expenditure_percent_of_GDP +
              Five_Year_GDP_Growth_Rate + Public_Debt_percent_of_GDP +
              Income_Tax_Rate + Tariff_Rate + GDP_per_Capita_PPP + 
              Inflation_rate, data = TRAIN)
summary(lm_full)

# Add Predictions to TEST
TEST2 = TEST %>%
  add_predictions(lm_full, var="pred0") %>%
  add_residuals(lm_full, var="resid0") %>%
  select(Country, EFS, pred0, resid0)

# Calculate MAE
MAE.func(TEST2$resid0)
```

```{r, warning=FALSE}
# Model 1: High-Correlation Variables
lm1 <- lm(EFS ~ GDP_per_Capita_PPP + Tariff_Rate + Inflation_rate + Tax_Burden_percent_of_GDP, data = TRAIN)

# Model 2: Government-related Metrics with Interaction
lm2 <- lm(EFS ~ (Govt_Expenditure_percent_of_GDP * FDI_Inflow_Millions * Corporate_Tax_Rate) +
           (Tax_Burden_percent_of_GDP * poly(FDI_Inflow_Millions, 2)) +
           Govt_Expenditure_percent_of_GDP * poly(Tariff_Rate, 2) +
           poly(GDP_per_Capita_PPP, 2) + Inflation_rate, data = TRAIN)

# Model 3: Economic Growth Factors Interaction
lm3 <- lm(EFS ~ (Govt_Expenditure_percent_of_GDP * Tax_Burden_percent_of_GDP + poly(GDP_Growth_Rate, 2)) +
           Govt_Expenditure_percent_of_GDP * FDI_Inflow_Millions * Tariff_Rate +
           poly(GDP_per_Capita_PPP, 2) + Inflation_rate + Corporate_Tax_Rate, data = TRAIN)

# Add Predictions to TEST
TEST2 = TEST %>%
  add_predictions(lm1, var="pred1") %>%
  add_predictions(lm2, var="pred2") %>%
  add_predictions(lm3, var="pred3") %>%
  add_residuals(lm1, var="resid1") %>%
  add_residuals(lm2, var="resid2") %>%
  add_residuals(lm3, var="resid3") %>%
  select(Country, EFS, pred1:resid3)

# Calculate MAE
MAE.func(TEST2$resid1)
MAE.func(TEST2$resid2)
MAE.func(TEST2$resid3)
```


```{r, warning=FALSE}
# Create Folds for Cross Validation
EFI_FOLD = crossv_kfold(EFI_DATA, k=20)

# Train Model for CV (Using model 3)
train.model.func=function(data){
  mod = lm(EFS ~ (Govt_Expenditure_percent_of_GDP * Tax_Burden_percent_of_GDP + poly(GDP_Growth_Rate, 2)) +
           Govt_Expenditure_percent_of_GDP * FDI_Inflow_Millions * Tariff_Rate +
           poly(GDP_per_Capita_PPP, 2) + Inflation_rate + Corporate_Tax_Rate, data = data)
  return(mod)
}

# Add Predictions and Residual results from Cros Validation
EFI_PREDICT = EFI_FOLD %>%
  mutate(tr.model=map(train,train.model.func)) %>%
  mutate(predict = map2(test,tr.model,~augment(.y,newdata=.x))) %>%
  select(predict) %>%
  unnest(cols=c(predict)) %>%
  select(Country, EFS, .fitted, .resid) %>%
  arrange(desc(EFS))

head(EFI_PREDICT, 10)

MAE.func(EFI_PREDICT$.resid)
```




```{r}
# Write CSV
write.csv(EFI_PREDICT, "EFI_PREDICT.csv", row.names = F)
```

Jins Models:

```{r}
jin_model1 <- lm(EFS ~ GDP_per_Capita_PPP + 
                    Tariff_Rate + 
                    GDP_Growth_Rate + 
                    Inflation_rate + 
                    Tax_Burden_percent_of_GDP + 
                    Govt_Expenditure_percent_of_GDP + 
                    Corporate_Tax_Rate,
                 data = TRAIN)
summary(jin_model1)

train.model.func.jin=function(data){
  mod1 = lm(EFS ~ GDP_per_Capita_PPP + 
                    Tariff_Rate + 
                    GDP_Growth_Rate + 
                    Inflation_rate + 
                    Tax_Burden_percent_of_GDP + 
                    Govt_Expenditure_percent_of_GDP + 
                    Corporate_Tax_Rate,
                 data = data)
  return(mod1)
  }

TEST_jin <- TEST %>%
  add_predictions(jin_model1, var = "Pred_Jin") %>%
  add_residuals(jin_model1, var = "Resid_Jin") %>%
  select(EFS, Pred_Jin, Resid_Jin)

MAE.func(TEST_jin$Resid_Jin)
```
```{r}
jin_model1 <- lm(EFS ~ GDP_per_Capita_PPP + 
                    Tariff_Rate + 
                    GDP_Growth_Rate + 
                    Inflation_rate + 
                    Tax_Burden_percent_of_GDP + 
                    Govt_Expenditure_percent_of_GDP + 
                    Corporate_Tax_Rate,
                 data = TRAIN)


train.model.func.jin=function(data,I,J,K,L,M,N,O){
  mod=lm(EFS ~ poly(GDP_per_Capita_PPP,I)
         +poly(Tariff_Rate,J) +
           poly(GDP_Growth_Rate, K) +
           poly(Inflation_rate, L) +
           poly(Tax_Burden_percent_of_GDP, M) + 
           poly(Govt_Expenditure_percent_of_GDP, N) +
           poly(Corporate_Tax_Rate, O)
         ,data=data)
  return(mod)
}

#Creating an MAE

# Function to train models and compute MAE for all combinations
train.and.evaluate.models <- function(data){
  
# List to store results
results <- list()
  
  # Nested loops to iterate over all combinations of polynomial degrees
  for(I in 1:4){
    for(J in 1:4){
      for(K in 1:4){
        for(L in 1:4){
          for(M in 1:4){
            for(N in 1:4){
              for(O in 1:4){
                
                # Skip combinations where all degrees are 1 (no polynomial terms)
                if(all(c(I, J, K, L, M, N, O) == 1)) next
                
                # Train model
                mod <- train.model.func.jin(data, I, J, K, L, M, N, O)
                
                # Make predictions
                predictions <- predict(mod, newdata=data)
                
                # Compute MAE
                mae <- MAE.func(data$EFS - predictions)
                
                # Store results
                results[[paste(I, J, K, L, M, N, O, sep="_")]] <- mae
              }
            }
          }
        }
      }
    }
  }
  
  return(results)
}

# Call the function with your dataset
results <- train.and.evaluate.models(TRAIN)

# Print the results
print(results)


library(tibble)

# Convert the list to a tibble
results_tibble <- as_tibble(results, .name_repair = "minimal")

# Print the tibble
print(results_tibble)

MAE.results <- results_tibble %>%
  gather("1_1_1_1_1_1_2":"4_4_4_4_4_4_4", key = "degrees", value = "MAE.values") %>%
  arrange(MAE.values)
```

Doing the same thing above to embrey's model

Polynomial degrees that minimizes MAE:
4_4_2_2_1_4_3

EMbrey's model 3:

train.model.func=function(data){
  mod = lm(EFS ~ (Govt_Expenditure_percent_of_GDP * Tax_Burden_percent_of_GDP
                  + poly(GDP_Growth_Rate, 2)) +
           Govt_Expenditure_percent_of_GDP * FDI_Inflow_Millions * Tariff_Rate +
           poly(GDP_per_Capita_PPP, 2) + 
             Inflation_rate + 
             Corporate_Tax_Rate, data = data)
  return(mod)
}


```{r}
#Creating a linear model with the minimizing polynomial variabkles and embrey's interaction variables
jin_model2 <- lm(EFS ~ poly(GDP_per_Capita_PPP, 4) + 
                    poly(Tariff_Rate, 4) + 
                    poly(GDP_Growth_Rate, 2) + 
                    poly(Inflation_rate, 2) + 
                    (Tax_Burden_percent_of_GDP) + 
                    poly(Govt_Expenditure_percent_of_GDP, 4) + 
                    poly(Corporate_Tax_Rate,3) + 
                   (Govt_Expenditure_percent_of_GDP * Tax_Burden_percent_of_GDP) +
                   (Govt_Expenditure_percent_of_GDP * FDI_Inflow_Millions * Tariff_Rate)
                 ,data = TRAIN)
#R-squared is pretty high
summary(jin_model2)

#Adding predictions and residuals for the above model
TEST_jin2 <- TEST %>%
  add_predictions(jin_model2, var = "Pred_Jin") %>%
  add_residuals(jin_model2, var = "Resid_Jin") %>%
  select(EFS, Pred_Jin, Resid_Jin)

#achieved the lowest MAE yet of: 3.642329
MAE.func(TEST_jin2$Resid_Jin)
```


#What if we do a stepwise selection on the model above?
```{r}
none_model = lm(EFS ~ 1, data = TRAIN)

MSE = (summary(jin_model2)$sigma)^2

step(none_model, scope=list(upper=jin_model2), scale=MSE)


jin_model3 = lm(formula = EFS ~ poly(GDP_per_Capita_PPP, 4) + poly(GDP_Growth_Rate, 
    2) + Govt_Expenditure_percent_of_GDP + poly(Inflation_rate, 
    2) + poly(Tariff_Rate, 4) + Tax_Burden_percent_of_GDP, data = TRAIN)

summary(jin_model3)

anova(jin_model2, jin_model3)

TEST_jin3 <- TEST %>%
  add_predictions(jin_model3, var = "Pred_Jin") %>%
  add_residuals(jin_model3, var = "Resid_Jin") %>%
  select(EFS, Pred_Jin, Resid_Jin)

#achieved the lowest MAE yet of: 3.642329
MAE.func(TEST_jin3$Resid_Jin)

#STEPWISE REGRESSION DOES NOT HELP, IT GOT RID OF ALL INTERACTION VARIABLES BUT HAVING THEM IS BETTER BUT WITH THE ANOVA TEST THERE IS NO STATISTICAL SIGNIFICANE THAT THE INTERACTION VARIABLES ARE HELPFUL IN PREDICTING, MAE increases by a lot so stepwise did not help
```













